DRAFT: Go Implementation of Techniques for Weighted Language Equivalence

### 2020-10-08 paper/sections/algorithms.tex
refined f1-tolerance test, working on paper

\section{Algorithms}
The first algorithm we implement to compute language equivalence, called \texttt{HKC},
is adapted from \cite{DBLP:journals/corr/Bonchi0K17}. 
The algorithm returns \texttt{true} $\iff \sqmap{v_1} = \sqmap{v_2}$.
It was first introduced by 
Bonchi and Pous in \cite{bonchi2013checking}.
The algorithm, extending the Hopcroft and Karp procedure 
\cite{hopcroft1971linear} with \textit{congruence closure}, is 
proven to be sound and complete \cite{DBLP:journals/corr/Bonchi0K17}.
It is defined for 
WAs over semirings, but in this implementation we are only 
considering fields, in particular 
the field of real numbers ($\K = \R$).
The problem of checking language equivalence 
has been proven undecidable for an arbitrary semiring, so termination 
may not always be guaranteed. However, it has been shown to be decidable
for a broad range of semirings, for example, all the complete and
distributive lattices.
\texttt{HKC} computes $v_1 \sim_l v_2$ for a given weighted automaton
$W = (X, t, o)$ and two vectors $v_1, v_2 \in \K^X$. 
by computing a congruence closure,
and it does so without linearizing the state space.

We compare \texttt{HKC} with an algorithm called 
\textit{Backwards Partition Refinement}, that we will call BPR for short.

Adapted from \cite{BONCHI201277}, 
\texttt{BPR}, or \textit{Backwards Partition Refinement}
is a fixed-point iterative method that, given an LWA
$L = (V, t, o)$, it computes the whole

computes a basis of the subspace of $V$
representing the complementary relation of $\llwb$ (we later show it to be the 
orthogonal complement in case $V$ is an inner product space). 
Another version of the algorithm is defined in the same work,
called \textit{Forward Partition Refinement}, which directly computes
a basis for $\llwb$ but is shown to be way less efficient than the backwards version.
The advantage of the backwards
It has been first introduced in \cite{boreale2009weighted}.
Our implementation is directly modeled on the algorithm shown in 
\cite{boreale2009weighted}, since we are fixing the field 
to $\R$ and computing the orthogonal complements instead of 
dual spaces and annihilators. In \cite{BONCHI201277}, the authors 
extended.

@@

The \texttt{BPR} algorithm starts from a relation $R_0$, that is the complement 
of the relation identifying vectors with equal weights.
It then incrementally computes the space of all states that are reachable from 
$R_0$ in a \textit{backwards} direction. Intuitively, "going backwards" means 
working with the transpose transitions functions $t_a^T$.

In the next sections we will compare execution of our implementation of the algorithms
\texttt{BPR} and \texttt{HKC} to verify correctness,
and to provide insight on runtime results.
Lemma \ref{lem:coincide}, introduced above, is key to our work. By stating that 
$\llwb$ coincides with $\sim_l$, we can confidently say that the two algorithms 
compute an answer for same the decision problem:

@@

%TODO
TODO costo computazionale HKC.

\texttt{BPR} has a cost of $O(n^4)$ operations 
to initially compute the largest linear weighted bisimulation,
which can be eventually reduced to $O(n^3)$ \cite{BONCHI201277}.
In our implementation, by initially computing a basis of the orthogonal complement of $\llwb$,
the cost of checking if two vectors are language equivalent is then reduced to the
cost of matrix multiplication ($O(n^2)$). \texttt{BPR} is a great choice when we
have to decide if a large number of vectors in a WA are language equivalent.

@@

\subsection{HKC Algorithm}
We give a pseudocode definition of the \texttt{HKC} procedure:

@@

We now recall the backwards algorithm for computing $\llwb$ defined in \cite{BONCHI201277}.
The algorithm is defined by the iterative method:
\begin{eqnarray}
  R_0 = \mker{o}^0, & \quad & R_{i+1} = R_i + \sum_{a \in A} t_a^T(R_i) \label{back} 
\end{eqnarray}
Where $\mker{o}^0$ is an annihilator.
The algorithm stops when $R_{j+1} = R_j$. An index $j \leq \dim(V)$ is 
guaranteed to exist, such that the algorithms terminates at step $j$.
It follows that $\llwb = R_j^0$.
Proof is available in section 4.2 of \cite{BONCHI201277}

>>>>

\section{Algorithms}
The first algorithm we implement to compute language equivalence, called \texttt{HKC},
is adapted from \cite{DBLP:journals/corr/Bonchi0K17}. 
The algorithm returns \texttt{true} $\iff \sqmap{v_1} = \sqmap{v_2}$.
It was first introduced by 
Bonchi and Pous in \cite{bonchi2013checking}.
The algorithm, extending the Hopcroft and Karp procedure 
\cite{hopcroft1971linear} with \textit{congruence closure}, is 
proven to be sound and complete.
It is defined for 
WAs over semirings, but in this implementation we are only 
considering fields, in particular 
the field of real numbers ($\K = \R$).
The problem of checking language equivalence 
has been proven undecidable for an arbitrary semiring, so termination 
may not always be guaranteed. However, it has been shown to be decidable
for a broad range of semirings, for example, all the complete and
distributive lattices.
\texttt{HKC} computes $v_1 \sim_l v_2$ for a given weighted automaton
$W = (X, t, o)$ and two vectors $v_1, v_2 \in \K^X$. 
by computing a congruence closure,
and it does so without linearizing the state space.

We compare \texttt{HKC} with an algorithm called 
\textit{Backwards Partition Refinement}, that we will call BPR for short. 
Adapted from \cite{BONCHI201277}, 
\texttt{BPR}
is a fixed-point iterative method that, given an LWA
$L = (V, t, o)$, computes a basis of the subspace of $V$
representing the complementary relation of $\llwb$ (we later show it to be the 
orthogonal complement in case $V$ is an inner product space). 
Another version of the algorithm is defined in the same work,
called \textit{Forward Partition Refinement}, which directly computes
a basis for $\llwb$ but is shown to be way less efficient than the backwards version.

Our implementation is directly modeled on the algorithm shown in 
\cite{boreale2009weighted}, since we are fixing weights on $\R$ and computing the orthogonal complements instead of 
dual spaces and annihilators.

@@

From Lemma \ref{lem:coincide}, it follows that given an LWA $L = (V,o,t)$ and a corresponding
basis of $\llwb$ computed by BPR, one can check language equivalence 
of two vectors in the state space, $v_1 \sim_l v_2 $,  by checking if $(v_1 - v_2) \in \mker{\llwb}$.
Therefore, we can say that BPR "\textit{minimizes}", or it computes the whole
binary linear relation $\llwb$, coinciding with $\sim_l$.

The \texttt{BPR} algorithm starts from the basis of a relation $R_0$, that is the complement 
of the relation identifying vectors with equal weights.
It then incrementally computes the space of all states that are reachable from 
$R_0$ in a \textit{backwards} direction. Intuitively, "going backwards" means 
working with the transpose transitions functions $t_a^T$.

\texttt{BPR} has a cost of $O(n^4)$ operations 
to initially compute the largest linear weighted bisimulation,
which can be eventually reduced to $O(n^3)$ \cite{BONCHI201277}.
In our implementation, by initially computing a basis of the orthogonal complement of $\llwb$,
the cost of checking if two vectors are language equivalent is then reduced to the
cost of matrix multiplication ($O(n^2)$). It follows that 
\texttt{BPR} is a great choice when we
have to decide if a large number of vectors in a WA are language equivalent.

In the next sections we will compare execution results of our implementation of the algorithms
\texttt{BPR} and \texttt{HKC} to verify correctness,
and to provide insight on runtime results.
Lemma \ref{lem:coincide}, introduced above, is key to our work. By stating that 
$\llwb$ coincides with $\sim_l$, we can confidently say that the two algorithms 
compute an answer for same the decision problem:

@@

%TODO
%TODO costo computazionale HKC.

@@

\subsection{HKC Algorithm}
We give the pseudocode definition of the \texttt{HKC} procedure from \cite{DBLP:journals/corr/Bonchi0K17}:

@@

We now recall the backwards algorithm for computing $\llwb$ defined in \cite{BONCHI201277}.
The algorithm is defined by the iterative method:
\begin{eqnarray}
  R_0 = \mker{o}^0, & \quad & R_{i+1} = R_i + \sum_{a \in A} t_a^T(R_i) \label{back} 
\end{eqnarray}
Where $\mker{o}^0$ is an annihilator.

\begin{note}
  Given two vector spaces $V_1, V_2$ we write $V_1 + V_2$ to 
  denote $\mspan{V_1 \cup V_2}$
\end{note}

The algorithm stops when $R_{j+1} = R_j$. An index $j \leq \dim(V)$ is 
guaranteed to exist, such that the algorithms stops at step $j$.
It follows that $\llwb = R_j^0$.
Proof is available in section 4.2 of \cite{BONCHI201277}

### 2020-10-08 paper/sections/conclusion.tex
refined f1-tolerance test, working on paper

%TODO 
TODO da fare

>>>>

During the testing phase, we have found that on randomly generated automata 
over the $\R$ field, the probability of BPR returning an empty basis of $\llwb$
increases together with the number of states in the automaton, and the number of
 symbols in the alphabet.
In a given linear weighted automaton $L = (V, o,t )$
over the $\R$ field and an alphabet $A$, the chance of BPR returning an empty basis also
increases together with the maximum weight in modulo of the transition matrices,
$\forall a \in A, \; \max |(t_a)_{ij}|$, for $i,j=1\hdots,n$, with 
$n = \dim(V)$. Authors of \cite{BONCHI201277} have confirmed
that it is normal for this to happen and that this fact is not due
to an implementation error.

Since the tests in this work rely heavily on a non-empty basis of $\llwb$ to generate
language equivalent vectors, the automata with an empty $\mker{\llwb}$ had to be ignored during the 
tests. This introduced substantial overhead as the chance of BPR returning an empty basis 
grew closer to 1, together with the various parameter of the automata: 
most of the computing time was spent on generating and running BPR on automata which did
not contain any language equivalent vectors, making tests on large automata not possible.

To provide better runtime results for the language equivalence problem, 
a topic worth further attention is the development of a semi-randomized method to 
generate structured large sized automata that have a low chance of BPR
returning an empty basis of $\llwb$.

Another idea for the future of this implementation, is the addition
of a \texttt{Semiring} interface type that would permit much more insightful analysis
on weighted automata, not only on the field of real numbers.

To further verify of the algorithms implemented in this work, 
and any additional algorithm that may be implemented, 
one could compare the results of this package with the results 
of the PAWS tool for the analysis of weighted systems \cite{konig2017paws}, developed 
at the University of Duisburg-Essen.

### 2020-10-08 paper/sections/implementation.tex
refined f1-tolerance test, working on paper

\subsection{Data Structures}
In this implementation, the data structure for representing weighted automata is a \texttt{struct} containing
the following attributes:
\begin{enumerate}
    \item An integer \texttt{Dim} representing the number of states $n$. 
    (note: states are finite and indexed on natural numbers $0, \hdots, n-1$)
    \item A slice of strings representing the \textit{alphabet} $A$.
    \item A map of strings as keys (the alphabet symbols) and dense $n \times n$ \texttt{float64} matrices as values,
    representing the set of transition functions.
    \item A dense \texttt{float64} vector, representing the linearization of the output function. 
    \item A \texttt{float64} value representing the tolerance to be used in numerical computations.
    \item An optionally \texttt{nil}, dense floating point matrix providing a basis for the orthogonal 
    complement of $\llwb$
\end{enumerate}
\begin{note}
    \textit{Slices} in Go are a convenient and efficient extension of the concept of arrays: 
    they provide an abstraction for indexed, variable length sequences of typed data, and 
    provide useful helper functions for creating, appending and selecting elements. 
\end{note}
The definition can be found in file \texttt{automata/automata.go}.
We then provide methods for reading an automaton from a text stream, applying transition
and output functions to vectors, and generating random automata with real and natural valued weights.

@@

\subsection{Implementation of BPR}
Given a WA $L$, at the first step of \texttt{BPR}, we set $R_0 = o$, 
with $o$ being the dense vector representing the output function 
of $L$, as seen in \cite{BONCHI201277}.
For each step $i = 0, \hdots, j+1$, with $j \leq \dim(L)$, the implemented algorithm:
\begin{enumerate}
    \item For each $a \in A$ computes $t_a^T(R_i)$ through matrix multiplication
    \item Concatenates those matrices $t_a^T(R_i)$ to $R_i$ in a resulting matrix $G$ 
    \item Instead of checking linear independence of the columns through Gaussian elimination,
    it computes $R_{i+1}$ as the orthonormal basis of the column space of $G$, through singular 
    value decomposition.
\end{enumerate}

@@

To compute the orthogonal complement of a vector subspace $W$, we
compute $W^\perp = \mker{A^T}$, where $A$ is the matrix whose column space 
is $W$. Proof is available in \cite{ila}.

>>>>

\subsection{Data Structures}
In this implementation, the data structure for representing weighted automata is a \texttt{struct}:

\captionof{figure}{Source code for the Weighted Automaton data structure, found in \texttt{automata/automata.go}}
\lstinputlisting[language=go]{../automata/automata.go}

%\begin{enumerate}
%    \item An integer \texttt{Dim} representing the number of states $n$. 
%    (note: states are finite and indexed on natural numbers $0, \hdots, n-1$)
%    \item A slice of strings representing the \textit{alphabet} $A$.
%    \item A map of strings as keys (the alphabet symbols) and dense $n \times n$ \texttt{float64} matrices as values,
%    representing the set of transition functions.
%    \item A dense \texttt{float64} vector, representing the linearization of the output function. 
%    \item A \texttt{float64} value representing the tolerance to be used in numerical computations.
%    \item An optionally \texttt{nil}, dense floating point matrix providing a basis for the orthogonal 
%    complement of $\llwb$
%\end{enumerate}
\begin{note}
    \textit{Slices} in Go are a convenient and efficient extension of the concept of arrays: 
    they provide an abstraction for indexed, variable length sequences of typed data, and 
    provide useful helper functions for creating, appending and selecting elements. 
\end{note}

We then provide methods for reading an automaton from a text stream, applying transitions
and output functions to vectors, and generating random automata with real and natural valued weights.

@@

\subsection{Implementation of BPR}
Given a WA $L$, at the first step of \texttt{BPR}, we set $R_0 = o$, 
with $o$ being the dense vector representing the output function 
of $L$, as seen in \cite{BONCHI201277}.
To compute $R_{i+1}$ at each step, the implemented BPR algorithm:
\begin{enumerate}
    \item Computes $t_a^T(R_i)$ through matrix multiplication for each $a \in A$
    \item Concatenates the resulting matrices to $R_i$ in a resulting matrix $G$ 
    \item Computes $R_{i+1}$ as the orthonormal basis of the column space of $G$, through singular 
    value decomposition.
\end{enumerate}

@@

To compute the orthogonal complement of a vector subspace $W$, we
compute $W^\perp = \mker{A^T}$, where $A$ is the matrix whose column space 
is $W$.

### 2020-10-08 paper/sections/introduction.tex
refined f1-tolerance test, working on paper

\section{Introduction}
\label{sec:intro}
TODO da fare

Checking whether two nondeterministic finite automata (NFA) accept the same
language is important in many application domains such as compiler construction and model checking. 
Unfortunately, solving this problem is costly: it is
PSPACE-complete (TODO cita) 
% TODO cita

In \cite{DBLP:journals/corr/Bonchi0K17}, up-to techniques are defined for
weighted systems over arbitrary semirings, while in \cite{BONCHI201277}, up-to techniques
are defined for Linear Weighted Automata (LWAs), under a more abstract coalgebraic perspective.

A well detailed example of the comparison of algorithms to compute 
language equivalence, precisely between \texttt{HKC} and an alternative
algorithm called the \textit{antichain algorithm} (\cite{de2006antichains}),
was published in 2017 \cite{fu2017equivalence}.

>>>>

\section{Introduction}
\label{sec:intro}

Weighted automata (WAs) are a generalization of non-deterministic automata.
When reading a symbol, a non-deterministic automaton 
can transition in different states simultaneously. Weighted automata
introduce \textit{weights} over transitions, which can for example, 
represent the cost of a transition, 
or in probabilistic systems, the chance of such transition to happen.
WAs can be represented with a set of states, an output function and
a set of transition matrices, indexed over the symbols in the alphabet
the automaton can read. While those automata are typically defined over semirings,
our implementation will focus for simplicity on automata with 
transitions defined over the real number field. 
The current configuration of a finite weighted automaton $W$,
defined on $n$ states, will be represented with a column vector of length
$n$, with values over the semiring or field on which transition weights of the
automaton $W$ are also defined.

The goal of this work is to provide an high-performance implementation 
in the Go programming language of two different techniques 
to compute \textit{weighted language equivalence}.
Such equivalence relation is a bisimulation:
a relation $R$ is a bisimulation whenever two states $v_1, v_2$ in $R$ 
can simulate each other, resulting in a pair that is still in $R$.
Two state vectors $v_1, v_2$ in a weighted automaton are said to be 
weighted language equivalent, written as $v_1 \sim_l v_2$, when 
they simulate each other by accepting the same words with the same resulting output weights.

The first technique we implement, 
defined in \cite{DBLP:journals/corr/Bonchi0K17}, is an up-to technique
for weighted language equivalence called HKC. It is defined for
weighted systems over arbitrary semirings and can be implemented with 
set theoretic constructs. The second technique is defined in \cite{BONCHI201277}:
a coalgebraic perspective is adopted to define a technique for language equivalence 
which exploits the linear representation of an automaton. 
This latter technique "\textit{minimizes}": by linearizing the state space of a
weighted automaton, 
it computes a basis for an entire linear relation (see definition \ref{def:linrel})
which coincides
with weighted language equivalence.
This technique for finite weighted automata over fields was first 
introduced by Michele Boreale in \cite{boreale2009weighted}.

Another example of the comparison between algorithms to compute 
language equivalence, precisely between HKC and an alternative
algorithm called the antichain algorithm (\cite{de2006antichains}),
was published in 2017 \cite{fu2017equivalence}.

### 2020-10-08 paper/sections/notation.tex
first public release
refined f1-tolerance test, working on paper

\begin{note}
  Given two vector spaces $V_1, V_2$ we write $V_1 + V_2$ to 
  denote $\mspan{V_1 \cup V_2}$
\end{note}

@@

Two vectors $v_1, v_2 \in \K^{X\times 1}$ are called weighted language equivalent, 
denoted with  $v_1 \sim_l v_2 $ if and only if 
$ \sqmap {v_1} = \sqmap{v_2}$. One can extend the notion of language 
equivalence to states rather than for vectors by assigning 
to each state $x \in X$ the corresponding  unit vector 
$e_x \in \K^X$. When given an initial state $i$ for a weighted automaton, 
the language  of the automaton can be defined as $\sqmap{i}$.

\begin{defn}
  A binary relation $R \subseteq X \times Y$ between two sets $X, Y$ is a 
  subset of the 
  cartesian product of the sets. A relation is called \textit{homogeneous} 
  or an \textit
  {endorelation} if it is a binary relation over $X$ and itself: 
  $R \subseteq X \times
   X$. 
  In such case, it is simply called a binary relation over $X$.
  An \textit{equivalence relation} is a binary relation that is reflexive, 
  symmetric and
  transitive. 
\end{defn}

@@

An equivalence relation which is compatible with all the operations of
the algebraic structure on which it is defined on, is called a 
\textit{congruence relation}. Compatibility with the algebraic structure operations
means that algebraic operations applied on equivalent elements will still
yield equivalent elements.

@@

\begin{defn}
  A \textit{linear weighted automaton} (in short, LWA) over the field $\K$ and 
  an alphabet $A$
  is a triple  $L = (V, o, \left\lbrace t_a \right\rbrace_{a \in A})$ 
  where $V$ is a vector space representing the state space, 
  $o: V \to \K$ is a linear map associating to each state its output weight,
  and $t = \left\lbrace t_a = V \times V \right\rbrace_{a \in A}$ is
  the set of transition functions, represented with liner maps 
  that for each input $a \in A$ associate the next state, in this case a vector
  in $V$.
  As in \cite{boreale2009weighted}, we have that $\dim{(L)} = \dim{(V)}$.
\end{defn}

Given a weighted automaton, one can build a corresponding linear weighted automaton
by considering the free vector space generated by the set of states $X$ in the WA,
and by linearizing $o$ and $t$. If $X$ is finite, as in our implementations
of the algorithms, we can use the same matrices for 
$t$ and $o$ in both the WA and the corresponding LWA.
We are only considering a finite number of states and therefore finite dimensional
vector spaces. Let $n$ be the number of states in an WA.
We have that in the corresponding LWA, the transition functions $t_a$ are still
 represented as
$\K^{n \times n}$ matrices. $o \in \K^{1 \times n}$ is represented as a row vector.
$t_a(v)$ denotes the vector obtained by multiplying the matrix $t_a$ by the column 
vector $v  \in \K^{n \times 1}$. $o(v)$ denotes the scalar $s \in \K$ obtained by 
dot product of the row vector $o$ with $v \in \K^{n \times 1}$.

\begin{defn}
  The language recognized by a vector $v \in V$ of an LWA $(V,o,t)$ is defined
  for all words $w \in A*$ as $\langlwa{v}{V}(w) = o(v_n)$ where $v_n$ is the 
  vector reached from $v$ through the composition of the transition functions
  corresponding to the words in $w$.

@@

We define $\llwb$ as the behavioral equivalence for a given LWA $(V, o, t)$ as

@@

Proof is available in section 3.3 of \cite{BONCHI201277}

@@

\begin{defn}
  \textit{Linear Relations:}\\
  Let $U$ be a subspace of $V$. The binary relation $R_U$ over $V$ is defined by
  \begin{equation*}
    \begin{aligned}
      v_1 \; R_U \; v_2 \quad \iff \quad v_1 - v_2 \in U
    \end{aligned}
  \end{equation*}
  The relation $R$ is linear if there exists a subspace $U$ such that $R \equiv R_U$.
  A linear relation is a total equivalence relation on $V$.
\end{defn}

@@

  Let $(X, o, t)$ be a WA and $(\K^X, o^\sharp, t^\sharp)$ the corresponding linear 
  weighted automaton. Then $\forall x \in X, \;\; \sqmap{x} = \langlwa{x}{\K^X}$
\end{lem}

>>>>

@@

Two vectors $v_1, v_2 \in \K^{X\times 1}$ are called weighted language equivalent, 
denoted with  $v_1 \sim_l v_2 $ if and only if 
$ \sqmap {v_1} = \sqmap{v_2}$. One can extend the notion of language 
equivalence to states rather than vectors by assigning 
to each state $x \in X$ the corresponding  unit vector 
$e_x \in \K^X$. When given an initial state $i$ for a weighted automaton, 
the language  of the automaton can be defined as $\sqmap{i}$.

\begin{defn}
  A binary relation $R \subseteq X \times Y$ between two sets $X, Y$ is a 
  subset of the 
  cartesian product of the sets. A relation is called \textit{homogeneous} 
  or an \textit
  {endorelation} if it is a binary relation over $X$ and itself: 
  $R \subseteq X \times
   X$. 
  In such case, it is simply called a binary relation over $X$.
  An \textit{equivalence relation} is a binary relation that is reflexive, 
  symmetric and
  transitive. 
  An equivalence relation which is compatible with all the operations of
the algebraic structure on which it is defined on, is called a 
\textit{congruence relation}. Compatibility with the algebraic structure operations
means that algebraic operations applied on equivalent elements will still
yield equivalent elements.

@@

\begin{defn}
  \label{def:congclos}
  \textit{Generating set for the congruence closure:} \\
  Let $\K$ be a field, $X$ a finite set and $R \subseteq \K^X \times \K^X$
  a relation. Let $(v, v') \in \K^X \times \K^X$ be a pair of vectors.
  The generating set is defined as  $U_R = \left\lbrace u - u' \mid (u, u') \in R \right\rbrace$.
  Then $(v, v') \in c(R) \iff (v - v') \in U_R$
\end{defn}

@@

\begin{defn}
  A \textit{linear weighted automaton} (in short, LWA) over the field $\K$ and 
  an alphabet $A$
  is a triple  $L = (V, o, \left\lbrace t_a \right\rbrace_{a \in A})$ 
  where $V$ is a vector space representing the state space, 
  $o: V \to \K$ is a linear map associating to each state its output weight,
  and $t = \left\lbrace t_a \in \K^{V \times V} \right\rbrace_{a \in A}$ is
  the set of transition functions, represented with liner maps 
  that for each input $a \in A$ associate the next state, in this case a vector
  in $V$.
  As seen in \cite{boreale2009weighted}, we have that $\dim{(L)} = \dim{(V)}$.
\end{defn}

Given a weighted automaton, one can build a corresponding linear weighted automaton
by considering the free vector space generated by the set of states $X$ in the WA,
and by linearizing $o$ and $t$. If $X$ is finite we can use the same matrices for 
$t$ and $o$ in both the WA and the corresponding LWA.
We are only considering a finite number of states and therefore finite dimensional
vector spaces. Let $n$ be the number of states in an WA.
We have that in the corresponding LWA, the transition functions $t_a$ are still
 represented as
$\K^{n \times n}$ matrices. $o \in \K^{1 \times n}$ is represented as a row vector.
$t_a(v)$ denotes the vector obtained by multiplying the matrix $t_a$ by the column 
vector $v  \in \K^{n \times 1}$. $o(v)$ denotes the scalar $s \in \K$ obtained by 
dot product of the row vector $o$ with $v \in \K^{n \times 1}$.

\begin{defn}
  The language recognized by a vector $v \in V$ in an LWA $(V,o,t)$ is defined
  for all words $w \in A*$ as $\langlwa{v}{V}(w) = o(v_n)$ where $v_n$ is the 
  vector reached from $v$ through the composition of the transition functions
  corresponding to each symbol in $w$.

@@

We define the equivalence relation  $\llwb$ for a given LWA $L = (V, o, t)$ as

@@

Proofs are available in section 3.3 of \cite{BONCHI201277}

@@

\begin{defn}
  \label{def:linrel}
  \textit{Linear Relations:}\\
  Let $U$ be a subspace of $V$. The binary relation $R_U$ over $V$ is defined by
  \begin{equation*}
    \begin{aligned}
      v_1 \; R_U \; v_2 \quad \iff \quad v_1 - v_2 \in U
    \end{aligned}
  \end{equation*}
  The relation $R$ is linear if there exists a subspace $U$ such that $R = R_U$.
  A linear relation is a total equivalence relation on $V$.
\end{defn}

@@

  %Let $(X, o, t)$ be a WA and $(\K^X, o^\sharp, t^\sharp)$ the corresponding linear 
  %weighted automaton. Then $\forall x \in X, \;\; \sqmap{x} = \langlwa{x}{\K^X}$
\end{lem}

\begin{figure}[htbp!]
  \centering
  \input{automata.tex}
  \begin{equation*}
    \begin{aligned}
      o = \begin{pmatrix}
        2 & 1 & 1
      \end{pmatrix} & \quad &  
      T_a = \begin{pmatrix}
        1 & \frac{1}{3} & 1 \\ 
        0 & 0 & 0 \\
        0 & 0 & 0 \\ 
      \end{pmatrix} & \quad &
      T_b = \begin{pmatrix}
        1 & 0 & 0 \\ 
        0 & 0 & 3 \\
        0 & \frac{1}{3} & 0 \\ 
      \end{pmatrix} & \quad &
    \end{aligned}
  \end{equation*}
  \caption{Example of a Weighted Automata}
  \label{fig:autexmp}
\end{figure}

### 2020-10-08 paper/sections/runtime.tex
fix
refined f1-tolerance test, working on paper

TODO da fare e sistemare.

The program was executed on an x86\_64 AMD Ryzen 5 2600X Six Core CPU, running at 3.6GHz with 32GB RAM, 
running Void Linux. 
For 3 test runs on 100000 automata composed of 4 states, 2 alphabet symbols and transition
matrices with weights on natural numbers $\leq 3$, with a tolerance of 10e-14,
we found a non empty kernel of $\llwb$ for an average of 1718 automata.

%TODO
(TODO spiega: più aumentano i simboli/stati più si riduce la probabilità di 
trovare $\llwb$ non vuoto?? investiga, trova un modo di generare automi con LLWB 
non vuoto)

By generating 1000 random couples of language equivalent vectors $(v_1, v_2)$ such that
$(v_1 - v_2) \in \mker{\llwb}$ and 1000 couples of completely random vectors
$(w_1, w_2)$, for each automata, we verified with 100\% confidence that:

\begin{center}
    $v_1 \; R_j \; v_2 \iff$ \texttt{HKC}$(v_1, v_2) =$ \texttt{false}
\end{center}

Where $R_j$ is the largest linear weighted bisimulation $\llwb$ computed by \texttt{BPR}.

%TODO

Sample output (19992 KBs in memory):
\begin{verbatim}
    $ time go run main.go 
    tests completed            100000...

    1000 samples per automaton
    100000 automata
    LLWP is not empty for 1726 automata
    language equivalence HKC = BPR for 3452000 samples
    BPR equivalence verified for 1726000 samples
    HKC equivalence verified for 1726000 samples
    confidence is 100%

    real    0m21.322s
    user    0m23.614s
    sys     0m0.961s
\end{verbatim}

>>>>

%TODO da fare e sistemare.

%\subsection{Tolerance and F1 Score}
The first runtime result we introduce is the study of the influence 
of the tolerance value choice over the correctness of the algorithms results.
Tests took place on randomly generated automata. A basis of  $\llwb$ was 
computed for each automaton with BPR. An arbitrary number of vector pairs were generated 
as a random linear combination of the vectors in the spanning set of $\llwb$. 
An identical quantity of completely random vector pairs were generated.
BPREquivalence and HKC procedures were then executed with those pairs in input.

By defining pairs of vectors as

\begin{itemize}
    \item \textit{true positives} (TP) if reported as language equivalent by both 
    procedures
    \item \textit{true negatives} (TN) if reported as not language equivalent by both 
    procedures
    \item \textit{false positives} (FP) if reported as language equivalent by HKC but not by BPR
    \item \textit{false negatives} (FN) if reported as  language equivalent by BPR but not by HKC
\end{itemize}

We have introduced \textit{accuracy, precision, recall} and \textit{F1 score} over 
the tests results:

\begin{equation*}
    \begin{aligned}
        \text{Accuracy} = \dfrac{TP + TN}{TP + TN + FP + FN} & \quad & 
        \text{Recall} = \dfrac{TP}{TP + FN} \\ && \\
        \text{Precision} = \dfrac{TP}{TP + FP} & \quad &
        \text{F1} = 2 * \dfrac{\text{Precision} \cdot \text{Recall}}{\text{Precision} + \text{Recall}}  \\ && \\
    \end{aligned}
\end{equation*}

We have then collected 
the F1 score in relation to varying tolerance values.
The plot is shown in Figure \ref{fig:f1}.  
The program was executed on an x86\_64 AMD Ryzen 5 2600X Six Core CPU, running at 3.6GHz with 32GB RAM, 
running Void Linux. Executed sequentially, F1 score tests took 2m26.19s
Parallelized with a worker pool, the tests took 1m13.12s.

% //TODO: condizionamento BPR
TODO: BPR conditioning

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.8\textwidth]{./plots/f1-tol.pdf}

    \caption{F1 score over tolerance tests, 1000 vector pairs tested per automaton.}
    \label{fig:f1}
\end{figure}

TODO empty LLWB chance, memory and time plots, varying automata size parameters
